\documentclass[11pt,letterpaper]{article}

% Load some basic packages that are useful to have
% and that should be part of any LaTeX installation.
%
% be able to include figures
\usepackage{graphicx}
% get nice colors
\usepackage{xcolor}

% change default font to Palatino (looks nicer!)
\usepackage[latin1]{inputenc}
\usepackage{mathpazo}
\usepackage[T1]{fontenc}
% load some useful math symbols/fonts
\usepackage{latexsym,amsfonts,amsmath,amssymb}

% comfort package to easily set margins
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{hyperref}
\usepackage[all]{hypcap}
% control some spacings
%
% spacing after a paragraph\begin{figure}[bth]
\setlength{\parskip}{.15cm}
% indentation at the top of a new paragraph
\setlength{\parindent}{0.0cm}

\begin{document}

\begin{center}
\Large
Ay190 -- Worksheet 09\\
David Vartanyan\\
Date: \today
\end{center}

\section{}

The goal is to solve the linear system of equations:

\begin{equation}
  A x = b
\end{equation}

for $nxn$ matrices $A$ and $nx1$ vectors $b$. 
I write a fanciful code that checks that each matrix $A$ is square and of same dimension as the vector $b$. I find the determinant, which is nonzero in every case. Thus these matrices can be solved.For matrices LSE4$\_$m and LSE5$\_$m I find the natural log of the determinant because attempting to calculate the determinant itself encounters overflow error.

The dimensions and determinants are available for any of the LSEs in ws9.py.

\begin{tabular}{|c|c|c|c|c|c|}
  \hline
& LSE1 & LSE2 & LSE3  & LSE4 & LSE5\\ \hline
Det Sign & 1 & -1 & -1 & -1 & 1\\
Det Natural Log & -4.53358731971 & 77.1114857221 & 226.07016741 & 1930.40191136 &  4563.75910742\\
Matrix Dim & 10 & 100 & 200 & 1000 & 2000\\ \hline
\end{tabular}

\section{}

I use the code available at \url{http://shiftycow.net/gauss-elimination-functions-in-python-and-matlab/}. Note this code does not pivot the matrices. It takes as an input an $n\,x\,n+1$ matrix whose $n+1$th column is the vector $b$. Code is available at ws9b1.py. I use time.clock to time the evaluation.

\begin{tabular}{|c|c|c|c|c|c|}
  \hline
& LSE1 & LSE2 & LSE3  & LSE4 & LSE5\\ \hline
time (s) & 0.0 & 0.08 & 0.3 & 21.42 & 173.52\\ \hline

\end{tabular}

\section{}

I used scipy.linalg to solve the system of equations using time.clock to time the results as above. Code is avaialable at ws9c.py.

\begin{tabular}{|c|c|c|c|c|c|}
  \hline
& LSE1 & LSE2 & LSE3  & LSE4 & LSE5\\ \hline
time (s) & 0.0 & 0.0 & 0.1& 0.24 & 1.78\\ \hline

\end{tabular}

Clearly scipy.linalg runs much faster (by two orders for the last three matrices), but we again see exponential growth in run time. Double the array dimension increases evaluation time by $9$ and not 4 as we would expect for an evaluation of time order $n^{2}$.

\end{document}


